---
title: "Concepts of Bayesian Data Analysis: Project4"
author:
  - Charles Muiruri^[Hasselt University, charles.muiruri@student.uhasselt.be]
  - Gianni Guarraci^[Hasselt University, gianni.guarraci@student.uhasselt.be]
  - Juan Vanegas Jadan^[Hasselt University, juan.vanegasjadan@student.uhasselt.be]
  - Ronald Makanga^[Hasselt University, ronald.makanga@student.uhasselt.be]
date: Report generated - `r format(Sys.time(), "%B %d, %Y")`
output: 
  pdf_document: 
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.height = 12,
  fig.width = 12
  )
```


```{r load_packages, include=FALSE}

if (!require(pacman)) install.packages("pacman")
p_load(rjags, coda, nimble, R2OpenBUGS, ggplot2, here, dplyr, ggpubr, tidyr, R2OpenBUGS)

```


```{r load_data, include=FALSE}
projdata <- as.data.frame(read.csv(here("data/projectdata.txt")))|>
  mutate(less_133pc = ifelse(Poverty == "<133% FPL", 1,0),
         btn133_400_pc = ifelse(Poverty == "133% to <400% FPL", 1,0),
         great_400pc = ifelse(Poverty == ">400% FPL", 1,0))
```


```{r prepDta, include=FALSE}
# Data prep for bugs
model_data <- list(
  Y = projdata$Vaccinated,
  N = projdata$Sample.Size,
  btn133_400_pc = projdata$btn133_400_pc,
  great_400pc = projdata$great_400pc,
  J = nrow(projdata)
)
```


```{r initial_values, include=FALSE}
model_inits <- list(
  list(beta0 = 0, beta1 = 0, beta2 = 0)
)

parameters <- c("beta0", "beta1", "beta2")

model_inits_ <- list(
  list(beta0 = 0, beta1 = 0, beta2 = 0),
  list(beta0 = 1, beta1 = 1, beta2 = 0),
  list(beta0 = 1, beta1 = 0, beta2 = 1)
)

```


```{r define_model, include=FALSE}
model1 <- function(){
  for (i in 1:J){
    Y[i] ~ dbin(p[i], N[i])
    logit(p[i]) <- beta0 + beta1*btn133_400_pc[i] + beta2*great_400pc[i]
  }
  #priors
  beta0 ~ dnorm(0, 0.001)
  beta1 ~ dnorm(0, 0.001)
  beta2 ~ dnorm(0, 0.001)
}

# Write model to file
write.model(model1, here("models/model1code.txt"))
# View file
# file.show(here("models/model1code.txt"))
```

\newpage 
# Modeling using Open Bugs

## Model

```{r model}
model.out <- bugs(model_data, model_inits, 
                  parameters = parameters, model.file = here("models/model1code.txt"),
                  n.chains = 1, n.iter = 10000, n.burnin = 5000, codaPkg = TRUE, debug = FALSE)
model.outq3 <- bugs(model_data, model_inits_, 
                  parameters = parameters, model.file = here("models/model1code.txt"),
                  n.chains = 3, n.iter = 10000, n.burnin = 2000, codaPkg = TRUE, debug = FALSE)

# debug=TRUE opens openBug and displays traceplots and summaries

out <- read.bugs(model.out)
out2 <- read.bugs(model.outq3)
# summary(out)
# HPDinterval(out, prob = 0.95)
```

## Question 1:

### The bugs model is as shown below

```
model
{
    for (i in 1:J) {
        Y[i] ~ dbin(p[i], N[i])
        logit(p[i]) <- beta0 + beta1 * btn133_400_pc[i] + beta2 * 
            great_400pc[i]
    }
    beta0 ~ dnorm(0.00000E+00, 0.001)
    beta1 ~ dnorm(0.00000E+00, 0.001)
    beta2 ~ dnorm(0.00000E+00, 0.001)
}

```



```{r Q1}
inverse_logit <- function(x) {
  1 / (1 + exp(-x))
}
out.summary <- summary(out)
beta0_mean <- out.summary[[1]]["beta0","Mean"]
beta1_mean <- out.summary[[1]]["beta1","Mean"]
beta2_mean <- out.summary[[1]]["beta2","Mean"]
pi_133 <- inverse_logit(beta0_mean)
pi_133_400 <- inverse_logit(beta1_mean + beta0_mean)
pi_400 <- inverse_logit(beta2_mean + beta0_mean)

cat("The mean probability of vaccination coverage for each poverty group is as follows:\n")
cat("<133% FPL: ", pi_133, "\n")
cat("133% to <400% FPL: ", pi_133_400, "\n")
cat(">400% FPL: ", pi_400, "\n")

## output should be the Bugs model

```


## Question 2

```{r}
parames <- c("beta0", "beta1", "beta2")

# Run MCMC
model_jag <- jags.model("model1code.txt", data = model_data, n.chains = 3, n.adapt = 1000)
update(model_jag, 1000)
samples_jag <- coda.samples(model_jag, variable.names = parames, n.iter = 10000)

# Check convergence
#summary(samples_jag)
#plot(samples_jag)
#gelman.diag(samples_jag)
#effectiveSize(samples_jag)


```

## Question 3: Density plots

```{r density_plots, fig.height=8, fig.width=8}
# Prep data for density and trace plots
mcmc_samples <- as.mcmc(out)

mcmc_sample2 <- bind_rows(lapply(out2, as.data.frame), .id="chain")|>
  pivot_longer(cols=-chain, names_to = "Parameter", values_to = "Value")


mcmc_df <- as.data.frame(mcmc_samples)
mcmc_df$iteration <- 1:nrow(mcmc_df)
mcmc_long_2 <- pivot_longer(mcmc_df, cols = -iteration, names_to = "Parameter", values_to = "Value")

hpd <- HPDinterval(out, prob = 0.95)
#plot(out)
points_data <- data.frame(x = inverse_logit(mcmc_df$beta0), x1 = inverse_logit(mcmc_df$beta0 + mcmc_df$beta1),x2 = inverse_logit(mcmc_df$beta0 + mcmc_df$beta2),
                          x3 = mcmc_df$deviance,y = rep(0, nrow(mcmc_df)))

# ggarrange(p_beta0 <- ggplot(mcmc_df, aes(x = beta0)) +
#   geom_density(fill = "blue", alpha = 0.1) +
#   geom_point(data = points_data, aes(x = x, y = y)) +
#   labs(title = "Posterior Distribution of beta0", x = "beta0", y = "Density")+
#   geom_vline(aes(xintercept = hpd[[1]]["beta0", "lower"]), linetype = "dashed", color = "red") +
#   geom_vline(aes(xintercept = hpd[[1]]["beta0", "upper"]), linetype = "dashed", color = "red") +
#   theme_bw(base_size = 12)+
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
# 
# p_beta1 <- ggplot(mcmc_df, aes(x = beta1)) +
#   geom_density(fill = "green", alpha = 0.1) +
#   geom_point(data = points_data, aes(x = x1, y = y)) +
#   labs(title = "Posterior Distribution of beta1", x = "beta1", y = "Density") +
#   geom_vline(aes(xintercept = hpd[[1]]["beta1", "lower"]), linetype = "dashed", color = "red") +
#   geom_vline(aes(xintercept = hpd[[1]]["beta1", "upper"]), linetype = "dashed", color = "red") +
#   theme_bw(base_size = 12)+
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
# 
# p_beta2 <- ggplot(mcmc_df, aes(x = beta2)) +
#   geom_density(fill = "red", alpha = 0.1) +
#   geom_point(data = points_data, aes(x = x2, y = y)) +
#   labs(title = "Posterior Distribution of beta2", x = "beta2", y = "Density")+
#   geom_vline(aes(xintercept = hpd[[1]]["beta2", "lower"]), linetype = "dashed", color = "red") +
#   geom_vline(aes(xintercept = hpd[[1]]["beta2", "upper"]), linetype = "dashed", color = "red") +
#   theme_bw(base_size = 12)+
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
# 
# p_dev <- ggplot(mcmc_df, aes(x = deviance)) +
#   geom_density(fill = "yellow", alpha = 0.1) +
#   geom_point(data = points_data, aes(x = x3, y = y)) +
#   labs(title = "Posterior Distribution of deviance", x = "Deviance", y = "Density")+
#   geom_vline(aes(xintercept = hpd[[1]]["deviance", "lower"]), linetype = "dashed", color = "red") +
#   geom_vline(aes(xintercept = hpd[[1]]["deviance", "upper"]), linetype = "dashed", color = "red") +
#   theme_bw(base_size = 12)+
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()), nrow = 2, ncol = 2)

mcmc_sample2 <- bind_rows(lapply(out2, as.data.frame), .id="chain")|>
  pivot_longer(cols=-chain, names_to = "Parameter", values_to = "Value")|>
  filter(Parameter != "deviance")

hpd_intervals2 <- HPDinterval(out2, prob = 0.95)
intvals <- data.frame(Parameter = c("beta0", "beta1", "beta2"),
                      Lower = c(mean(hpd_intervals2[[1]]["beta0", "lower"], hpd_intervals2[[2]]["beta0", "lower"],hpd_intervals2[[3]]["beta0", "lower"]), mean(hpd_intervals2[[1]]["beta1", "lower"], hpd_intervals2[[2]]["beta1", "lower"], hpd_intervals2[[3]]["beta1", "lower"]), mean(hpd_intervals2[[1]]["beta2", "lower"], hpd_intervals2[[2]]["beta2", "lower"], hpd_intervals2[[3]]["beta2", "lower"])),
                      Upper = c(mean(hpd_intervals2[[1]]["beta0", "upper"], hpd_intervals2[[2]]["beta0", "upper"], hpd_intervals2[[3]]["beta0", "upper"]), mean(hpd_intervals2[[1]]["beta1", "upper"], hpd_intervals2[[2]]["beta1", "upper"], hpd_intervals2[[3]]["beta1", "upper"]), mean(hpd_intervals2[[1]]["beta2", "upper"], hpd_intervals2[[2]]["beta2", "upper"], hpd_intervals2[[3]]["beta2", "upper"])))

mcmc_sample2fin <- left_join(mcmc_sample2, intvals, by = "Parameter")#|>
  # pivot_longer(cols = c("Value", "Lower", "Upper"), names_to = "Type", values_to = "Value")

# mcmcfinb0 <- mcmc_sample2fin|>filter(Parameter == "beta0")

# mcmc_sample2fin2 <- mcmc_sample2fin|>
#   mutate(Value=case_when(Parameter == "beta0" ~ plogis(Value),
#                          Parameter == "beta1"  ~ plogis(Value + mcmcfinb0$Value[1]),
#                          Parameter == "beta2" ~ plogis(Value + mcmcfinb0$Value[1]),
#                          TRUE ~ Value))|>
#   pivot_wider(names_from = Type, values_from = Value)

# mcmc_sample2_2 <- mcmc_sample2 %>%
#   mutate(
#     Estimate = case_when(
#       Parameter == "beta0" ~ Value,
#       Parameter == "beta1" ~ Value + get("beta0", envir = .GlobalEnv),
#       Parameter == "beta2" ~ Value + get("beta0", envir = .GlobalEnv),
#       TRUE ~ Value
#     )
#   )


ggplot(mcmc_sample2fin, aes(x = Value, fill = chain)) +
  geom_density(alpha = 0.6) +
  facet_wrap(~ Parameter, scales = "free", nrow = 2) +
  geom_vline(aes(xintercept = Lower), linetype = "dashed", color = "red")+
  geom_vline(aes(xintercept = Upper), linetype = "dashed", color = "red") +
  labs(x = "Value", y = "Density", title = "Density Plot of MCMC Samples") +
  theme_minimal()

#summary(out)
#summary(samples_jag) ## no deviance
```


### Summary of posterior distribution

  - The Posterior density with relation to **$\beta_0$** has a mean of **`r plogis(summary(out2)[[1]]["beta0","Mean"])`** which is the probability of children vaccinated from households whose income is less than 133%,  with a 95% credible HPD interval of **(`r plogis(intvals[1, 2])`, `r plogis(intvals[1, 3])`)** in which the proportion lies with 95% probability.
  - The Posterior density with relation to **$\beta_1$** has a mean of **`r plogis(summary(out2)[[1]]["beta1","Mean"] + summary(out2)[[1]]["beta0","Mean"])`** which is the probability of children vaccinated from households whose income between 133% and 400% ,  with a 95% credible HPD interval of **(`r plogis(intvals[1, 2] + intvals[2, 2])`, `r plogis(intvals[1, 3] + intvals[2, 3])`)** in which the proportion lies with 95% probability.
  - The Posterior density with relation to **$\beta_2$** has a mean of **`r plogis(summary(out2)[[1]]["beta2","Mean"] + summary(out2)[[1]]["beta0","Mean"])`** which is the probability of children vaccinated from households whose income is more than 400%,  with a 95% credible HPD interval of**(`r plogis(intvals[1, 2] + intvals[3, 2])`, `r plogis(intvals[1, 3] + intvals[3, 3])`)** in which the proportion lies with 95% probability.




```{r trace_plots, eval=FALSE}

### Trace Plots
ggplot(mcmc_long|>filter(iteration > 4800), aes(x = iteration, y = Value, color = Parameter)) +
  geom_line() +
  scale_color_manual(values = c("blue", "green", "red", "yellow")) +
  facet_wrap(~ Parameter, scales = "free_y") +
  labs(title = "Trace Plots of MCMC Samples", x = "Iteration", y = "Parameter Value") +
  theme_bw(base_size = 12)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none")
```



```{r correlation_plot, fig.height=6, fig.width=7, eval=FALSE}

### Correlation plot
crosscorr.plot(out)

```



```{r autocorrelation_plot, fig.height=6, fig.width=7, eval=FALSE}
### Autocorrelation plot
autocorr.plot(out)
```


## QUESTION 4: Calculate the posterior probability that the vaccination coverage target (90%) is reached for each poverty group.


```{r Q4_calc, fig.height=6, fig.width=7}

# mcmc_samples contains the posterior samples
mcmc_samples <- as.mcmc(out)
posterior_samples <- as.data.frame(mcmc_samples)

# Function to calculate probability from log-odds
logit_to_prob <- function(logit) {
  exp(logit) / (1 + exp(logit))
}

# Calculate the probabilities for each poverty group
posterior_samples <- posterior_samples %>%
  mutate(
    pi_133less = logit_to_prob(beta0),
    pi_133_400 = logit_to_prob(beta0 + beta1),
    pi_400more = logit_to_prob(beta0 + beta2)
  )

# Calculate the proportion of samples where vaccination probability is >= 90%
target_coverage <- 0.90

prob_133less_reached <- mean(posterior_samples$pi_133less >= target_coverage)
prob_133_400_reached <- mean(posterior_samples$pi_133_400 >= target_coverage)
prob_400more_reached <- mean(posterior_samples$pi_400more >= target_coverage)

# Print the results
cat("Posterior probability that vaccination coverage target (90%) is reached:\n")
cat("<133% FPL: ", prob_133less_reached, "\n")
cat("133-400% FPL: ", prob_133_400_reached, "\n")
cat(">400% FPL: ", prob_400more_reached, "\n")

```

```{r, fig.height=6, fig.width=7}
# Combine the results into a data frame for plotting
posterior_probs_df <- data.frame(
  PovertyGroup = c("<133% FPL", "133-400% FPL", ">400% FPL"),
  Probability = c(prob_133less_reached, prob_133_400_reached, prob_400more_reached)
)

# Print the probabilities
posterior_probs_df
```

```{r, fig.height=6, fig.width=7}
# Plot the probabilities using ggplot2
ggplot(posterior_probs_df, aes(x = PovertyGroup, y = Probability)) +
  geom_bar(stat = "identity", fill = c("blue", "green", "red"), alpha = 0.7) +
  geom_text(aes(label = scales::percent(Probability, accuracy = 0.1)), vjust = -0.3, size = 5) +
  ylim(0, 1) +
  labs(
    title = "Posterior Probability of Reaching 90% Vaccination Coverage",
    x = "Poverty Group",
    y = "Probability"
  ) +
  theme_minimal(base_size = 15)

```

```{r Q4_plot, fig.height=6, fig.width=7}
# Assuming posterior_samples already contains the calculated probabilities
posterior_samples <- posterior_samples %>%
  mutate(
    pi_133less = logit_to_prob(beta0),
    pi_133_400 = logit_to_prob(beta0 + beta1),
    pi_400more = logit_to_prob(beta0 + beta2)
  )

# Function to create density plot with 90% threshold line
create_density_plot <- function(data, var, group_name) {
  ggplot(data, aes_string(x = var)) +
    geom_density(fill = "blue", alpha = 0.3) +
    geom_vline(xintercept = 0.9, linetype = "dashed", color = "red") +
    labs(title = paste("Posterior Distribution for", group_name),
         x = "Vaccination Probability",
         y = "Density") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
}

# Create density plots for each group
p1 <- create_density_plot(posterior_samples, "pi_133less", "<133% FPL")
p2 <- create_density_plot(posterior_samples, "pi_133_400", "133-400% FPL")
p3 <- create_density_plot(posterior_samples, "pi_400more", ">400% FPL")

# Arrange the plots
library(gridExtra)
grid.arrange(p1, p2, p3, ncol = 1)

```


## Question 5

```{r}

jags_posterior_samples <- as.data.frame(as.matrix(samples_jag))
# Calculate coverage reduction
coverage_reduction <- apply(jags_posterior_samples, 1, function(params) {
  beta1 <- params["beta1"]
  beta2 <- params["beta2"]

  reduction <- plogis(beta1) - plogis(beta2)
  return(reduction)
})

# Summary of coverage reduction
summary(coverage_reduction)
```


## Question 6 

Investigate whether the vaccination coverages are distinct at the different locations by adding a location-specific intercept.

For this, beta0 is assumed to be the logit of the proportion of vaccinated individuals for each location in the group less than 133% FPL, so we need to calculate it. So we will find just the proportion for groups 

```{r Q6, eval=FALSE}
model1 <- function(){
  for (i in 1:J){
    Y[i] ~ dbin(p[i], N[i])
    logit(p[i]) <- beta0[i] + beta1*btn133_400_pc[i] + beta2*great_400pc[i]
  }
  beta1 ~ dnorm(0, 0.001)
  beta2 ~ dnorm(0, 0.001)
}

write.model(model1, here("models/model2code.txt"))
projdata.beta0 <- projdata %>%
  filter(less_133pc == 1) %>%
  group_by(Geography) %>%
  mutate(beta0 = logit(sum(Vaccinated)/sum(Sample.Size))) %>%
  select(Geography, beta0)

projdata <- left_join(projdata, projdata.beta0, by = "Geography")

model_data_loc <- list(
  Y = projdata$Vaccinated,
  N = projdata$Sample.Size,
  beta0 = projdata$beta0,
  btn133_400_pc = projdata$btn133_400_pc,
  great_400pc = projdata$great_400pc,
  J = nrow(projdata),
  Location = projdata$Location
)

model_inits_loc <- list( list(beta1 = 0, beta2 = 0) )
parameters_loc <- c("beta1", "beta2")
model.out <- bugs(model_data_loc, model_inits_loc, 
                  parameters = parameters_loc, model.file = here("models/model2code.txt"),
                  n.chains = 1, n.iter = 10000, n.burnin = 5000, codaPkg = TRUE, debug = FALSE)

out.loc <- read.bugs(model.out)
```


```{r Q6_summary, eval=FALSE}

out.summary <- summary(out.loc)
beta1_mean <- out.summary[[1]]["beta1","Mean"]
beta2_mean <- out.summary[[1]]["beta2","Mean"]
pi_133 <- mean(inverse_logit(projdata$beta0))
pi_133_400 <- inverse_logit(beta1_mean + beta0_mean)
pi_400 <- inverse_logit(beta2_mean + beta0_mean)

cat("The mean probability of vaccination coverage for each poverty group is as follows:\n")
cat("Giving a value to beta0 means that the proportion for the group less than 133% FPL is the logit of beta0 \n")
cat("<133% FPL: ", pi_133, "\n")
cat("133% to <400% FPL: ", pi_133_400, "\n")
cat(">400% FPL: ", pi_400, "\n")
```


```{r question_6_jags}
projdata_jag <- projdata %>%
  mutate(location = as.numeric(factor(Geography)))

model_data_jags <- list(
  Y = projdata_jag$Vaccinated,
  N = projdata_jag$Sample.Size,
  btn133_400_pc = projdata_jag$btn133_400_pc,
  great_400pc = projdata_jag$great_400pc,
  location = projdata_jag$location,
  J = nrow(projdata_jag),
  J_locations = length(unique(projdata_jag$location))
)

modeljags <- function(){
  for (i in 1:J){
    Y[i] ~ dbin(p[i], N[i])
    logit(p[i]) <-  beta1*btn133_400_pc[i] + beta2*great_400pc[i] + beta_location[location[i]]
  }
  #priors
  beta0 ~ dnorm(0, 0.001)
  beta1 ~ dnorm(0, 0.001)
  beta2 ~ dnorm(0, 0.001)
  for (j in 1:J_locations) {
    beta_location[j] ~ dnorm(0, 0.001)
  }
}


params_loc_jags <- c("beta1", "beta2", "beta_location")
# Write model to file
write.model(modeljags, here("model3code.txt"))
# View file
#file.show(here("model3code.txt"))
jags_mod2 <- jags.model("model3code.txt", data = model_data_jags, n.chains = 3, n.adapt = 1000)
update(jags_mod2, 2000)
jags_samples_extended <- coda.samples(jags_mod2, variable.names = params_loc_jags, n.iter = 10000)

# Check convergence
#print(summary(jags_samples_extended))
plot(jags_samples_extended)
#gelman.diag(jags_samples_extended)
#effectiveSize(jags_samples_extended)

```

## Question 7


```{r}
jag7samples <- coda.samples(jags_mod2, variable.names = c("beta0", "beta1", "beta2", "beta_location"), n.iter = 10000)

# Convert to a data frame
#jag7posterior_samples <- do.call(rbind, lapply(jag7_samples_extended, as.data.frame))
posterior_samples2 <- as.matrix(jags_samples_extended)

#  Calculate the mean and 95% credible intervals
coverage_estimates <- apply(posterior_samples2, 2, function(x) {
  mean_val <- mean(x)
  ci <- quantile(x, c(0.025, 0.975))
  c(mean = mean_val, lower = ci[1], upper = ci[2])
})

coverage_df <- data.frame(location = rep(colnames(coverage_estimates), each = 3),
                          estimate_type = rep(c("mean", "lower.2.5%", "upper.97.5%"), times = ncol(coverage_estimates)),
                          coverage = as.vector(coverage_estimates),
                          stringsAsFactors = FALSE)

betas_loc <- coverage_df |>
  slice(1:6)

# logit_to_prob <- function(logit) {
#   exp(logit) / (1 + exp(logit))
# }

betas_loc_est <- coverage_df |>
  slice(7:nrow(coverage_df))|>
  mutate(coverage_133pc = plogis(coverage),
         coverage_133_400pc = case_when(
           estimate_type == "mean" ~ plogis(coverage + betas_loc$coverage[1]),
           estimate_type == "lower.2.5%" ~ plogis(coverage + betas_loc$coverage[2]),
           estimate_type == "upper.97.5%" ~ plogis(coverage + betas_loc$coverage[3])
         ),
         coverage_400pc = case_when(
           estimate_type == "mean" ~ plogis(coverage + betas_loc$coverage[4]),
           estimate_type == "lower.2.5%" ~ plogis(coverage + betas_loc$coverage[5]),
           estimate_type == "upper.97.5%" ~ plogis(coverage + betas_loc$coverage[6])
         ))|>select(-coverage)|>pivot_longer(cols = c(coverage_133pc, coverage_133_400pc, coverage_400pc), names_to = "povlevel", values_to = "coverage")|>pivot_wider(names_from = estimate_type, values_from = coverage)|>
  mutate(location2 = paste0(location, povlevel),
         location2 = factor(location2, levels = location2[order(-mean)]))

coverage_df_wide <- coverage_df %>%
  pivot_wider(names_from = estimate_type, values_from = coverage)

#  Plot the data

ggplot(betas_loc_est, aes(x = location2, y = mean, ymin = `lower.2.5%`, ymax = `upper.97.5%`)) +
  geom_pointrange() +
  coord_flip() +
  labs(x = "Location", y = "Estimated Vaccination Coverage")+
  theme_bw(base_size = 12)


```


## Question 8

```{r}

mississippi_index <- 3


jag7posterior_samples <- do.call(rbind, lapply(jags_samples_extended, as.data.frame))

mean_coverage_133_400FPL <- mean(plogis( jag7posterior_samples$beta1 + jag7posterior_samples[, grep("^beta_location", colnames(jag7posterior_samples))][, mississippi_index]))
mean_coverage_400FPL <- mean(plogis( jag7posterior_samples$beta2 + jag7posterior_samples[, grep("^beta_location", colnames(jag7posterior_samples))][, mississippi_index]))
mean_coverage_133FPL <- mean(plogis( jag7posterior_samples[, grep("^beta_location", colnames(jag7posterior_samples))][, mississippi_index]))

# Predict the number of vaccinated children
n_children <- 300
n_133_400FPL <- 100
n_400FPL <- 100
n_others <- 100

pred_vaccinated_133_400FPL <- n_133_400FPL * mean_coverage_133_400FPL
pred_vaccinated_400FPL <- n_400FPL * mean_coverage_400FPL
pred_vaccinated_133FPL <- n_others * mean_coverage_133FPL




datsr<-rbind(c("133-400% FPL", as.numeric(round(pred_vaccinated_133_400FPL,0))), c(">400% FPL", as.numeric(round(pred_vaccinated_400FPL,0))), c("<133% FPL", as.numeric(round(pred_vaccinated_133FPL,0))))

dfe<-datsr|>data.table::as.data.table()|>data.table::setnames(c("Poverty", "Estimate"))|>dplyr::mutate(Estimate = as.numeric(Estimate))|>janitor::adorn_totals(c("row"))
View(dfe)
dfe|>flextable::flextable()|>flextable::autofit()

```


\newpage

# Appendix
`R Code`
```

if (!require(pacman)) install.packages("pacman")
p_load(rjags, coda, nimble, R2OpenBUGS, ggplot2, here, dplyr, ggpubr, tidyr)

projdata <- as.data.frame(read.csv(here("data/projectdata.txt")))|>
  mutate(less_133pc = ifelse(Poverty == "<133% FPL", 1,0),
         btn133_400_pc = ifelse(Poverty == "133% to <400% FPL", 1,0),
         great_400pc = ifelse(Poverty == ">400% FPL", 1,0))
# Data prep for bugs
model_data <- list(
  Y = projdata$Vaccinated,
  N = projdata$Sample.Size,
  btn133_400_pc = projdata$btn133_400_pc,
  great_400pc = projdata$great_400pc,
  J = nrow(projdata)
)

model_inits <- list(
  list(beta0 = 0, beta1 = 0, beta2 = 0)
)

parameters <- c("beta0", "beta1", "beta2")

model1 <- function(){
  for (i in 1:J){
    Y[i] ~ dbin(p[i], N[i])
    logit(p[i]) <- beta0 + beta1*btn133_400_pc[i] + beta2*great_400pc[i]
  }
  #priors
  beta0 ~ dnorm(0, 0.001)
  beta1 ~ dnorm(0, 0.001)
  beta2 ~ dnorm(0, 0.001)
}

# Write model to file
write.model(model1, here("models/model1code.txt"))
# View file
file.show(here("models/model1code.txt"))

model.out <- bugs(model_data, model_inits, 
                  parameters = parameters, model.file = here("models/model1code.txt"),
                  n.chains = 1, n.iter = 10000, n.burnin = 5000, codaPkg = TRUE, 
                  debug = FALSE)

# debug=TRUE opens openBug and displays traceplots and summaries

# Model output
out <- read.bugs(model.out)
summary(out)

# Prep data for density and trace plots
mcmc_samples <- as.mcmc(out)
mcmc_df <- as.data.frame(mcmc_samples)
mcmc_df$iteration <- 1:nrow(mcmc_df)
mcmc_long <- pivot_longer(mcmc_df, cols = -iteration, names_to = "Parameter",
                          values_to = "Value")

points_data <- data.frame(x = mcmc_df$beta0, x1 = mcmc_df$beta1,x2 = mcmc_df$beta2,
                          x3 = mcmc_df$deviance,y = rep(0, nrow(mcmc_df)))

# Density plots
ggarrange(p_beta0 <- ggplot(mcmc_df, aes(x = beta0)) +
            geom_density(fill = "blue", alpha = 0.1) +
            geom_point(data = points_data, aes(x = x, y = y)) +
            labs(title = "Posterior Distribution of beta0", x = "beta0", y = "Density")+
            theme_bw(base_size = 12)+
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
          
          p_beta1 <- ggplot(mcmc_df, aes(x = beta1)) +
            geom_density(fill = "green", alpha = 0.1) +
            geom_point(data = points_data, aes(x = x1, y = y)) +
            labs(title = "Posterior Distribution of beta1", x = "beta1", y = "Density") +
            theme_bw(base_size = 12)+
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
          
          p_beta2 <- ggplot(mcmc_df, aes(x = beta2)) +
            geom_density(fill = "red", alpha = 0.1) +
            geom_point(data = points_data, aes(x = x2, y = y)) +
            labs(title = "Posterior Distribution of beta2", x = "beta2", y = "Density")+
            theme_bw(base_size = 12)+
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
          
          p_dev <- ggplot(mcmc_df, aes(x = deviance)) +
            geom_density(fill = "yellow", alpha = 0.1) +
            geom_point(data = points_data, aes(x = x3, y = y)) +
            labs(title = "Posterior Distribution of deviance", x = "Deviance", y = "Density")+
            theme_bw(base_size = 12)+
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()), 
          nrow = 2, ncol = 2)

# Trace plots
ggplot(mcmc_long, aes(x = iteration, y = Value, color = Parameter)) +
  geom_line() +
  scale_color_manual(values = c("blue", "green", "red", "yellow")) +
  facet_wrap(~ Parameter, scales = "free_y") +
  labs(title = "Trace Plots of MCMC Samples", x = "Iteration", y = "Parameter Value") +
  theme_bw(base_size = 12)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        legend.position = "none")

# Autocorrelation and crosscorrelation plots
crosscorr.plot(out)
autocorr.plot(out)

```


