---
output: 
  pdf_document: 
    latex_engine: xelatex
header-includes:
- \usepackage{graphics}
- \usepackage{graphicx}
- \usepackage{lscape}
- \usepackage{amsfonts}
- \usepackage{amsmath}
- \usepackage{amsthm}
- \usepackage{amssymb}
- \usepackage{latexsym}
- \usepackage{color}
- \usepackage{verbatim}
- \usepackage{fancyhdr}
- \usepackage{fancybox}
- \usepackage{titlesec}
- \pagenumbering{roman}
---

\newcommand{\groupname}{Introduction to Bayesian Inference - Project4}

\pagestyle{fancy}
\pagenumbering{roman}

\begin{center}
	% details
	\Large{
	Project: Introduction to Bayesian Inference (3579) \\
	Project 4 \\
	2023-2024 \\[0.5cm]
	1st year Master of Statistics \\
	Hasselt University	
	}
\end{center}

\vspace*{1cm}
\textbf{\large{Group members:}}   
 Ronald Makanga (2365466)  
 Juan Vanegas Jadan (2366405)  
 Gianni Guarraci (2366293)  
 Charles Muiruri (2365043)  

\vspace*{0.5cm}
\noindent\textit{Submission Date:} 01 June, 2024

\vspace*{2.5cm}
\textbf{\large{Lecturers:}}  
 Prof. Dr. Christel FAES
 
\newpage

\tableofcontents

\newpage

\pagenumbering{arabic}
\chead{\groupname}
\rhead{}
\lhead{}
\cfoot{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\setcounter{page}{1}

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.height = 12,
  fig.width = 12
  )
```


```{r load_packages, include=FALSE}

if (!require(pacman)) install.packages("pacman")
p_load(rjags, coda, nimble, R2OpenBUGS, ggplot2, here, dplyr, ggpubr, tidyr, R2OpenBUGS)

```


```{r load_data, include=FALSE}
projdata <- as.data.frame(read.csv(here("data/projectdata.txt")))|>
  mutate(less_133pc = ifelse(Poverty == "<133% FPL", 1,0),
         btn133_400_pc = ifelse(Poverty == "133% to <400% FPL", 1,0),
         great_400pc = ifelse(Poverty == ">400% FPL", 1,0))
```


```{r prepDta, include=FALSE}
# Data prep for bugs
model_data <- list(
  Y = projdata$Vaccinated,
  N = projdata$Sample.Size,
  btn133_400_pc = projdata$btn133_400_pc,
  great_400pc = projdata$great_400pc,
  J = nrow(projdata)
)
```


```{r initial_values, include=FALSE}
parameters <- c("beta0", "beta1", "beta2")

model_inits_ <- list(
  list(beta0 = 0, beta1 = 0, beta2 = 0),
  list(beta0 = 1, beta1 = 1, beta2 = 0),
  list(beta0 = 1, beta1 = 0, beta2 = 1)
)

```


```{r define_model, include=FALSE}
model1 <- function(){
  for (i in 1:J){
    Y[i] ~ dbin(p[i], N[i])
    logit(p[i]) <- beta0 + beta1*btn133_400_pc[i] + beta2*great_400pc[i]
  }
  #priors
  beta0 ~ dnorm(0, 0.001)
  beta1 ~ dnorm(0, 0.001)
  beta2 ~ dnorm(0, 0.001)
}

write.model(model1, here("models/model1code.txt"))
```

\newpage 
# Modeling using Open Bugs

## Model

```{r model}
suppressMessages({
  suppressWarnings({
model.outq3 <- bugs(model_data, model_inits, 
                  parameters = parameters, model.file = here("models/model1code.txt"),
                  n.chains = 3, n.iter = 10000, n.burnin = 2000, codaPkg = TRUE, debug = FALSE)

# debug=TRUE opens openBug and displays traceplots and summaries

out2 <- read.bugs(model.outq3)

  })
})
```

## Question 1:

### The bugs model is as shown below

```
model
{
    for (i in 1:J) {
        Y[i] ~ dbin(p[i], N[i])
        logit(p[i]) <- beta0 + beta1 * btn133_400_pc[i] + beta2 * 
            great_400pc[i]
    }
    beta0 ~ dnorm(0.00000E+00, 0.001)
    beta1 ~ dnorm(0.00000E+00, 0.001)
    beta2 ~ dnorm(0.00000E+00, 0.001)
}

```



```{r Q1}
inverse_logit <- function(x) {
  1 / (1 + exp(-x))
}
out.summary <- summary(out)
beta0_mean <- out.summary[[1]]["beta0","Mean"]
beta1_mean <- out.summary[[1]]["beta1","Mean"]
beta2_mean <- out.summary[[1]]["beta2","Mean"]
pi_133 <- inverse_logit(beta0_mean)
pi_133_400 <- inverse_logit(beta1_mean + beta0_mean)
pi_400 <- inverse_logit(beta2_mean + beta0_mean)

cat("The mean probability of vaccination coverage for each poverty group is as follows:\n")
cat("<133% FPL: ", pi_133, "\n")
cat("133% to <400% FPL: ", pi_133_400, "\n")
cat(">400% FPL: ", pi_400, "\n")


```


## Question 2

```{r}
parames <- c("beta0", "beta1", "beta2")

# Run MCMC
suppressMessages({
  suppressWarnings({
    model_jag <- jags.model("model1code.txt", data = model_data, n.chains = 3)
    samples_jag <- coda.samples(model_jag, variable.names = parames, n.iter = 10000)
  })
})



```

## MCMC method and check converge of the MCMC chains Convergence tests
In our analysis, we used Gibbs Sampling to estimate the posterior distributions for our three  unknown parameters (beta0, beta1 and beta2). To test for convergence for our MCMC chains, we used both Gelman-Rubin convergence diagnostic and trace plots. 

The Gelman-Rubin convergence diagnostic method allows us to compare within and between chain variances for each variable. Best results are obtained for parameters whose marginal posterior 
densities are approximately normal. To run the Gelman-Rubin convergence test, we set three MCMC 
chains and for each chain we set distinct starting values for our unknown parameters as shown in 
the code. We ran 10,000 MCMC trials with a burn-in value of 2,000 and a thining value of 1 for each chain. 

Then we produced gelman-rubin statistics and plot using the gelman.diag and gelman.plot 
functions in R. The gelman.diag function gives us the scale reduction factors for each 
parameter (beta0,beta1,and beta2). A factor of 1 means that the between variance and within 
chain variance are equal, larger values mean that there is still a notable difference between
chains. on the other hand, gelman.plot shows if the shrink factor has really converged, or 
whether it still fluctuating. Both are results from both our functions. 


```{r}
gelman.diag(samples_jag)
gelman.plot(samples_jag,ask=FALSE)

```

In addition, we also explored convergence of MCMC chains by looking at our trace plots for each of our unknown parameter. The trace plot shows the parameter value at the time t against the iteration
number. 

```{r}
par(mfrow=c(2,2))
traceplot(samples_jag)
```


## Question 3 

```{r density_plots, fig.height=8, fig.width=8}
# Prep data for density and trace plots
mcmc_samples <- as.mcmc(out)

mcmc_sample2 <- bind_rows(lapply(out2, as.data.frame), .id="chain")|>
  pivot_longer(cols=-chain, names_to = "Parameter", values_to = "Value")


mcmc_df <- as.data.frame(mcmc_samples)
mcmc_df$iteration <- 1:nrow(mcmc_df)
mcmc_long_2 <- pivot_longer(mcmc_df, cols = -iteration, names_to = "Parameter", values_to = "Value")

hpd <- HPDinterval(out, prob = 0.95)
#plot(out)
points_data <- data.frame(x = inverse_logit(mcmc_df$beta0), x1 = inverse_logit(mcmc_df$beta0 + mcmc_df$beta1),x2 = inverse_logit(mcmc_df$beta0 + mcmc_df$beta2),
                          x3 = mcmc_df$deviance,y = rep(0, nrow(mcmc_df)))

# ggarrange(p_beta0 <- ggplot(mcmc_df, aes(x = beta0)) +
#   geom_density(fill = "blue", alpha = 0.1) +
#   geom_point(data = points_data, aes(x = x, y = y)) +
#   labs(title = "Posterior Distribution of beta0", x = "beta0", y = "Density")+
#   geom_vline(aes(xintercept = hpd[[1]]["beta0", "lower"]), linetype = "dashed", color = "red") +
#   geom_vline(aes(xintercept = hpd[[1]]["beta0", "upper"]), linetype = "dashed", color = "red") +
#   theme_bw(base_size = 12)+
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
# 
# p_beta1 <- ggplot(mcmc_df, aes(x = beta1)) +
#   geom_density(fill = "green", alpha = 0.1) +
#   geom_point(data = points_data, aes(x = x1, y = y)) +
#   labs(title = "Posterior Distribution of beta1", x = "beta1", y = "Density") +
#   geom_vline(aes(xintercept = hpd[[1]]["beta1", "lower"]), linetype = "dashed", color = "red") +
#   geom_vline(aes(xintercept = hpd[[1]]["beta1", "upper"]), linetype = "dashed", color = "red") +
#   theme_bw(base_size = 12)+
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
# 
# p_beta2 <- ggplot(mcmc_df, aes(x = beta2)) +
#   geom_density(fill = "red", alpha = 0.1) +
#   geom_point(data = points_data, aes(x = x2, y = y)) +
#   labs(title = "Posterior Distribution of beta2", x = "beta2", y = "Density")+
#   geom_vline(aes(xintercept = hpd[[1]]["beta2", "lower"]), linetype = "dashed", color = "red") +
#   geom_vline(aes(xintercept = hpd[[1]]["beta2", "upper"]), linetype = "dashed", color = "red") +
#   theme_bw(base_size = 12)+
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
# 
# p_dev <- ggplot(mcmc_df, aes(x = deviance)) +
#   geom_density(fill = "yellow", alpha = 0.1) +
#   geom_point(data = points_data, aes(x = x3, y = y)) +
#   labs(title = "Posterior Distribution of deviance", x = "Deviance", y = "Density")+
#   geom_vline(aes(xintercept = hpd[[1]]["deviance", "lower"]), linetype = "dashed", color = "red") +
#   geom_vline(aes(xintercept = hpd[[1]]["deviance", "upper"]), linetype = "dashed", color = "red") +
#   theme_bw(base_size = 12)+
#   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()), nrow = 2, ncol = 2)

mcmc_sample2 <- bind_rows(lapply(out2, as.data.frame), .id="chain")|>
  pivot_longer(cols=-chain, names_to = "Parameter", values_to = "Value")|>
  filter(Parameter != "deviance")

hpd_intervals2 <- HPDinterval(out2, prob = 0.95)
intvals <- data.frame(Parameter = c("beta0", "beta1", "beta2"),
                      Lower = c(mean(hpd_intervals2[[1]]["beta0", "lower"], hpd_intervals2[[2]]["beta0", "lower"],hpd_intervals2[[3]]["beta0", "lower"]), mean(hpd_intervals2[[1]]["beta1", "lower"], hpd_intervals2[[2]]["beta1", "lower"], hpd_intervals2[[3]]["beta1", "lower"]), mean(hpd_intervals2[[1]]["beta2", "lower"], hpd_intervals2[[2]]["beta2", "lower"], hpd_intervals2[[3]]["beta2", "lower"])),
                      Upper = c(mean(hpd_intervals2[[1]]["beta0", "upper"], hpd_intervals2[[2]]["beta0", "upper"], hpd_intervals2[[3]]["beta0", "upper"]), mean(hpd_intervals2[[1]]["beta1", "upper"], hpd_intervals2[[2]]["beta1", "upper"], hpd_intervals2[[3]]["beta1", "upper"]), mean(hpd_intervals2[[1]]["beta2", "upper"], hpd_intervals2[[2]]["beta2", "upper"], hpd_intervals2[[3]]["beta2", "upper"])))

mcmc_sample2fin <- left_join(mcmc_sample2, intvals, by = "Parameter")#|>
  # pivot_longer(cols = c("Value", "Lower", "Upper"), names_to = "Type", values_to = "Value")

# mcmcfinb0 <- mcmc_sample2fin|>filter(Parameter == "beta0")

# mcmc_sample2fin2 <- mcmc_sample2fin|>
#   mutate(Value=case_when(Parameter == "beta0" ~ plogis(Value),
#                          Parameter == "beta1"  ~ plogis(Value + mcmcfinb0$Value[1]),
#                          Parameter == "beta2" ~ plogis(Value + mcmcfinb0$Value[1]),
#                          TRUE ~ Value))|>
#   pivot_wider(names_from = Type, values_from = Value)

# mcmc_sample2_2 <- mcmc_sample2 %>%
#   mutate(
#     Estimate = case_when(
#       Parameter == "beta0" ~ Value,
#       Parameter == "beta1" ~ Value + get("beta0", envir = .GlobalEnv),
#       Parameter == "beta2" ~ Value + get("beta0", envir = .GlobalEnv),
#       TRUE ~ Value
#     )
#   )


ggplot(mcmc_sample2fin, aes(x = Value, fill = chain)) +
  geom_density(alpha = 0.6) +
  facet_wrap(~ Parameter, scales = "free", nrow = 2) +
  geom_vline(aes(xintercept = Lower), linetype = "dashed", color = "red")+
  geom_vline(aes(xintercept = Upper), linetype = "dashed", color = "red") +
  labs(x = "Value", y = "Density", title = "Density Plot of MCMC Samples") +
  theme_minimal()

```


### Summary of posterior distribution

  - The Posterior density with relation to **$\beta_0$** has a mean of **`r plogis(summary(out2)[[1]]["beta0","Mean"])`** which is the probability of children vaccinated from households whose income is less than 133%,  with a 95% credible HPD interval of **(`r plogis(intvals[1, 2])`, `r plogis(intvals[1, 3])`)** in which the proportion lies with 95% probability.
  - The Posterior density with relation to **$\beta_1$** has a mean of **`r plogis(summary(out2)[[1]]["beta1","Mean"] + summary(out2)[[1]]["beta0","Mean"])`** which is the probability of children vaccinated from households whose income between 133% and 400% ,  with a 95% credible HPD interval of **(`r plogis(intvals[1, 2] + intvals[2, 2])`, `r plogis(intvals[1, 3] + intvals[2, 3])`)** in which the proportion lies with 95% probability.
  - The Posterior density with relation to **$\beta_2$** has a mean of **`r plogis(summary(out2)[[1]]["beta2","Mean"] + summary(out2)[[1]]["beta0","Mean"])`** which is the probability of children vaccinated from households whose income is more than 400%,  with a 95% credible HPD interval of**(`r plogis(intvals[1, 2] + intvals[3, 2])`, `r plogis(intvals[1, 3] + intvals[3, 3])`)** in which the proportion lies with 95% probability.

## Question 4


```{r Q4_calc, fig.height=6, fig.width=7}

# mcmc_samples contains the posterior samples
mcmc_samples <- as.mcmc(out)
posterior_samples <- as.data.frame(mcmc_samples)

# Function to calculate probability from log-odds
logit_to_prob <- function(logit) {
  exp(logit) / (1 + exp(logit))
}

# Calculate the probabilities for each poverty group
posterior_samples <- posterior_samples %>%
  mutate(
    pi_133less = logit_to_prob(beta0),
    pi_133_400 = logit_to_prob(beta0 + beta1),
    pi_400more = logit_to_prob(beta0 + beta2)
  )

# Calculate the proportion of samples where vaccination probability is >= 90%
target_coverage <- 0.90

prob_133less_reached <- mean(posterior_samples$pi_133less >= target_coverage)
prob_133_400_reached <- mean(posterior_samples$pi_133_400 >= target_coverage)
prob_400more_reached <- mean(posterior_samples$pi_400more >= target_coverage)

# Print the results
cat("Posterior probability that vaccination coverage target (90%) is reached:\n")
cat("<133% FPL: ", prob_133less_reached, "\n")
cat("133-400% FPL: ", prob_133_400_reached, "\n")
cat(">400% FPL: ", prob_400more_reached, "\n")

```

```{r, fig.height=6, fig.width=7}
# Combine the results into a data frame for plotting
posterior_probs_df <- data.frame(
  PovertyGroup = c("<133% FPL", "133-400% FPL", ">400% FPL"),
  Probability = c(prob_133less_reached, prob_133_400_reached, prob_400more_reached)
)|>mutate(PovertyGroup = factor(PovertyGroup, levels = c("<133% FPL", "133-400% FPL", ">400% FPL")))

```


`r posterior_probs_df`


```{r, fig.height=6, fig.width=8}
# Plot the probabilities using ggplot2
ggplot(posterior_probs_df, aes(x = PovertyGroup, y = Probability)) +
  geom_bar(stat = "identity", fill = c("blue", "green", "red"), alpha = 0.7) +
  geom_text(aes(label = scales::percent(Probability, accuracy = 0.1)), vjust = -0.3, size = 5) +
  ylim(0, 1) +
  labs(
    title = "Posterior Probability of Reaching 90% Vaccination Coverage",
    x = "Poverty Group",
    y = "Probability"
  ) +
  theme_minimal(base_size = 15)

```



```{r Q4_plot, fig.height=6, fig.width=8}
# Assuming posterior_samples already contains the calculated probabilities
posterior_samples <- posterior_samples %>%
  mutate(
    pi_133less = logit_to_prob(beta0),
    pi_133_400 = logit_to_prob(beta0 + beta1),
    pi_400more = logit_to_prob(beta0 + beta2)
  )

# Function to create density plot with 90% threshold line
create_density_plot <- function(data, var, group_name) {
  ggplot(data, aes_string(x = var)) +
    geom_density(fill = "blue", alpha = 0.3) +
    geom_vline(xintercept = 0.9, linetype = "dashed", color = "red") +
    labs(title = paste("Posterior Distribution for", group_name),
         x = "Vaccination Probability",
         y = "Density") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
}

# Create density plots for each group
p1 <- create_density_plot(posterior_samples, "pi_133less", "<133% FPL")
p2 <- create_density_plot(posterior_samples, "pi_133_400", "133-400% FPL")
p3 <- create_density_plot(posterior_samples, "pi_400more", ">400% FPL")

# Arrange the plots
library(gridExtra)
grid.arrange(p1, p2, p3, ncol = 2)

```


## Question 5

```{r}
posterior_samples <- posterior_samples %>%
  mutate(
    diff_133_400_vs_133less = pi_133_400 - pi_133less,
    diff_400more_vs_133less = pi_400more - pi_133less,
    diff_400more_vs_133_400 = pi_400more - pi_133_400
  )

# Function to create density plot for differences
create_diff_density_plot <- function(data, var, group_comparison) {
  ggplot(data, aes_string(x = var)) +
    geom_density(fill = "blue", alpha = 0.3) +
    labs(title = stringr::str_wrap(paste("Posterior Distribution of Difference in Vaccination Probability", group_comparison), width = 60),
         x = "Difference in Vaccination Probability",
         y = "Density") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
}

# Create density plots for differences in vaccination probabilities
p1 <- create_diff_density_plot(posterior_samples, "diff_133_400_vs_133less", "133-400% FPL vs <133% FPL")
p2 <- create_diff_density_plot(posterior_samples, "diff_400more_vs_133less", ">400% FPL vs <133% FPL")
p3 <- create_diff_density_plot(posterior_samples, "diff_400more_vs_133_400", ">400% FPL vs 133-400% FPL")


library(gridExtra)
grid.arrange(p1, p2, p3, ncol = 2)
```
The mean reduction in vaccination coverage between the poverty groups is as follows:
- The mean reduction in vaccination coverage between the poverty groups less than 133% FPL and 133% to 400% FPL is `r mean(posterior_samples$diff_133_400_vs_133less)`.
- The mean reduction in vaccination coverage between the poverty groups less than 133% FPL and greater than 400% FPL is `r mean(posterior_samples$diff_400more_vs_133less)`.
- The mean reduction in vaccination coverage between the poverty groups 133% to 400% FPL and greater than 400% FPL is `r mean(posterior_samples$diff_400more_vs_133_400)`.

## Question 6 

Investigate whether the vaccination coverages are distinct at the different locations by adding a location-specific intercept.



```{r question_6_jags}
projdata_jag <- projdata %>%
  mutate(location = as.numeric(factor(Geography)))

model_data_jags <- list(
  Y = projdata_jag$Vaccinated,
  N = projdata_jag$Sample.Size,
  btn133_400_pc = projdata_jag$btn133_400_pc,
  great_400pc = projdata_jag$great_400pc,
  location = projdata_jag$location,
  J = nrow(projdata_jag),
  J_locations = length(unique(projdata_jag$location))
)

modeljags <- function(){
  for (i in 1:J){
    Y[i] ~ dbin(p[i], N[i])
    logit(p[i]) <-  beta1*btn133_400_pc[i] + beta2*great_400pc[i] + beta_location[location[i]]
  }
  #priors
  beta1 ~ dnorm(0, 0.001)
  beta2 ~ dnorm(0, 0.001)
  for (j in 1:J_locations) {
    beta_location[j] ~ dnorm(0, 0.001)
  }
}


model.inits2 <- list(
  list(beta_location = rep(1, model_data_jags$J_locations), beta1 = 0, beta2 = 0),
  list(beta_location = rep(1, model_data_jags$J_locations), beta1 = 1, beta2 = 0),
  list(beta_location = rep(1, model_data_jags$J_locations), beta1 = 0, beta2 = 1)
)


params_loc_jags <- c("beta1", "beta2", "beta_location")

# Write model to file
write.model(modeljags, here("model3code.txt"))
suppressMessages({
  suppressWarnings({
    jags_mod2 <- jags.model("model3code.txt", data = model_data_jags,inits = model.inits2, n.chains = 3, n.adapt = 1000)
    update(jags_mod2, 2000)
    jags_samples_extended <- coda.samples(jags_mod2, variable.names = params_loc_jags, n.iter = 10000)
  })
})

```

### The bugs model is as shown below

```
model
{
    for (i in 1:J) {
        Y[i] ~ dbin(p[i], N[i])
        logit(p[i]) <- beta1 * btn133_400_pc[i] + beta2 * great_400pc[i] + 
            beta_location[location[i]]
    }
    beta1 ~ dnorm(0.00000E+00, 0.001)
    beta2 ~ dnorm(0.00000E+00, 0.001)
    for (j in 1:J_locations) {
        beta_location[j] ~ dnorm(0.00000E+00, 0.001)
    }
}

```

### Gelman-Rubin Convergence Diagnostic

```{r}

gelman.diag(jags_samples_extended)
gelman.plot(jags_samples_extended,ask=FALSE)

```

## Question 7

```{r}

posterior_samples2 <- as.matrix(jags_samples_extended)

#  Calculate the mean and 95% credible intervals
coverage_estimates <- apply(posterior_samples2, 2, function(x) {
  mean_val <- mean(x)
  ci <- quantile(x, c(0.025, 0.975))
  c(mean = mean_val, lower = ci[1], upper = ci[2])
})

coverage_df <- data.frame(location = rep(colnames(coverage_estimates), each = 3),
                          estimate_type = rep(c("mean", "lower.2.5%", "upper.97.5%"), times = ncol(coverage_estimates)),
                          coverage = as.vector(coverage_estimates),
                          stringsAsFactors = FALSE)

betas_loc <- coverage_df |>
  slice(1:6)

betas_loc_est <- coverage_df |>
  slice(7:nrow(coverage_df))|>
  mutate(coverage_133pc = plogis(coverage),
         coverage_133_400pc = case_when(
           estimate_type == "mean" ~ plogis(coverage + betas_loc$coverage[1]),
           estimate_type == "lower.2.5%" ~ plogis(coverage + betas_loc$coverage[2]),
           estimate_type == "upper.97.5%" ~ plogis(coverage + betas_loc$coverage[3])
         ),
         coverage_400pc = case_when(
           estimate_type == "mean" ~ plogis(coverage + betas_loc$coverage[4]),
           estimate_type == "lower.2.5%" ~ plogis(coverage + betas_loc$coverage[5]),
           estimate_type == "upper.97.5%" ~ plogis(coverage + betas_loc$coverage[6])
         ))|>select(-coverage)|>pivot_longer(cols = c(coverage_133pc, coverage_133_400pc, coverage_400pc), names_to = "povlevel", values_to = "coverage")|>pivot_wider(names_from = estimate_type, values_from = coverage)|>
  mutate(
         location = factor(location, levels = c("beta_location[1]", "beta_location[2]", "beta_location[3]", "beta_location[4]", "beta_location[5]", "beta_location[6]", "beta_location[7]", "beta_location[8]", "beta_location[9]"), labels = c("Florida", "Georgia", "Mississippi", "Missouri", "North Carolina", "Oklahoma", "Pennsylvania", "Washington", "Wisconsin")),
         povlevel = factor(povlevel, levels = c("coverage_133pc", "coverage_133_400pc", "coverage_400pc"), labels = c("<133% FPL", "133% to <400% FPL", ">400% FPL")),
         location2 = paste0(location,"_", povlevel),
         location2 = factor(location2, levels = location2[order(-mean)]))

#  Plot the data

ggplot(betas_loc_est, aes(x = location2, y = mean, ymin = `lower.2.5%`, ymax = `upper.97.5%`, colour = povlevel)) +
  geom_pointrange() +
  coord_flip() +
  labs(x = "Geographic location", y = "Estimated Vaccination Coverage", colour = "Poverty Level")+
  theme_bw(base_size = 12)

```
This caterpillar plot displays the expected coverage for each geographic region within our data. The point represents the meancoverage while the error bars represent the 95% credible interval. The plot shows that the expected coverage for each region is distinct.

## Question 8

```{r}

mississippi_index <- 3


jag7posterior_samples <- do.call(rbind, lapply(jags_samples_extended, as.data.frame))

mean_coverage_133_400FPL <- mean(plogis( jag7posterior_samples$beta1 + jag7posterior_samples[, grep("^beta_location", colnames(jag7posterior_samples))][, mississippi_index]))
mean_coverage_400FPL <- mean(plogis( jag7posterior_samples$beta2 + jag7posterior_samples[, grep("^beta_location", colnames(jag7posterior_samples))][, mississippi_index]))
mean_coverage_133FPL <- mean(plogis( jag7posterior_samples[, grep("^beta_location", colnames(jag7posterior_samples))][, mississippi_index]))

# Predict the number of vaccinated children
n_children <- 300
n_133_400FPL <- 100
n_400FPL <- 100
n_others <- 100

pred_vaccinated_133_400FPL <- n_133_400FPL * mean_coverage_133_400FPL
pred_vaccinated_400FPL <- n_400FPL * mean_coverage_400FPL
pred_vaccinated_133FPL <- n_others * mean_coverage_133FPL




datsr<-rbind(c("133-400% FPL", as.numeric(round(pred_vaccinated_133_400FPL,0))), c(">400% FPL", as.numeric(round(pred_vaccinated_400FPL,0))), c("<133% FPL", as.numeric(round(pred_vaccinated_133FPL,0))))

dfe<-datsr|>data.table::as.data.table()|>data.table::setnames(c("Poverty", "Estimate"))|>dplyr::mutate(Estimate = as.numeric(Estimate))|>janitor::adorn_totals(c("row"))
View(dfe)
dfe|>flextable::flextable()|>flextable::autofit()

```
This table shows the estimated number of vaccinated children in Mississippi by poverty group. The estimated number of vaccinated children from households with income less than 133% FPL is `r round(pred_vaccinated_133FPL,0)`, from households with income between 133% and 400% FPL is  `r round(pred_vaccinated_133_400FPL,0)`,and from households with income greater than 400% FPL is `r round(pred_vaccinated_400FPL,0)`.

\newpage

# Appendix
`R Code`
```

if (!require(pacman)) install.packages("pacman")
p_load(rjags, coda, nimble, R2OpenBUGS, ggplot2, here, dplyr, ggpubr, tidyr)

projdata <- as.data.frame(read.csv(here("data/projectdata.txt")))|>
  mutate(less_133pc = ifelse(Poverty == "<133% FPL", 1,0),
         btn133_400_pc = ifelse(Poverty == "133% to <400% FPL", 1,0),
         great_400pc = ifelse(Poverty == ">400% FPL", 1,0))
# Data prep for bugs
model_data <- list(
  Y = projdata$Vaccinated,
  N = projdata$Sample.Size,
  btn133_400_pc = projdata$btn133_400_pc,
  great_400pc = projdata$great_400pc,
  J = nrow(projdata)
)

model_inits <- list(
  list(beta0 = 0, beta1 = 0, beta2 = 0)
)

parameters <- c("beta0", "beta1", "beta2")

model1 <- function(){
  for (i in 1:J){
    Y[i] ~ dbin(p[i], N[i])
    logit(p[i]) <- beta0 + beta1*btn133_400_pc[i] + beta2*great_400pc[i]
  }
  #priors
  beta0 ~ dnorm(0, 0.001)
  beta1 ~ dnorm(0, 0.001)
  beta2 ~ dnorm(0, 0.001)
}

# Write model to file
write.model(model1, here("models/model1code.txt"))
# View file
file.show(here("models/model1code.txt"))

model.out <- bugs(model_data, model_inits, 
                  parameters = parameters, model.file = here("models/model1code.txt"),
                  n.chains = 1, n.iter = 10000, n.burnin = 5000, codaPkg = TRUE, 
                  debug = FALSE)

# debug=TRUE opens openBug and displays traceplots and summaries

# Model output
out <- read.bugs(model.out)
summary(out)

# Prep data for density and trace plots
mcmc_samples <- as.mcmc(out)
mcmc_df <- as.data.frame(mcmc_samples)
mcmc_df$iteration <- 1:nrow(mcmc_df)
mcmc_long <- pivot_longer(mcmc_df, cols = -iteration, names_to = "Parameter",
                          values_to = "Value")

points_data <- data.frame(x = mcmc_df$beta0, x1 = mcmc_df$beta1,x2 = mcmc_df$beta2,
                          x3 = mcmc_df$deviance,y = rep(0, nrow(mcmc_df)))

# Density plots
ggarrange(p_beta0 <- ggplot(mcmc_df, aes(x = beta0)) +
            geom_density(fill = "blue", alpha = 0.1) +
            geom_point(data = points_data, aes(x = x, y = y)) +
            labs(title = "Posterior Distribution of beta0", x = "beta0", y = "Density")+
            theme_bw(base_size = 12)+
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
          
          p_beta1 <- ggplot(mcmc_df, aes(x = beta1)) +
            geom_density(fill = "green", alpha = 0.1) +
            geom_point(data = points_data, aes(x = x1, y = y)) +
            labs(title = "Posterior Distribution of beta1", x = "beta1", y = "Density") +
            theme_bw(base_size = 12)+
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
          
          p_beta2 <- ggplot(mcmc_df, aes(x = beta2)) +
            geom_density(fill = "red", alpha = 0.1) +
            geom_point(data = points_data, aes(x = x2, y = y)) +
            labs(title = "Posterior Distribution of beta2", x = "beta2", y = "Density")+
            theme_bw(base_size = 12)+
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()),
          
          p_dev <- ggplot(mcmc_df, aes(x = deviance)) +
            geom_density(fill = "yellow", alpha = 0.1) +
            geom_point(data = points_data, aes(x = x3, y = y)) +
            labs(title = "Posterior Distribution of deviance", x = "Deviance", y = "Density")+
            theme_bw(base_size = 12)+
            theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()), 
          nrow = 2, ncol = 2)

# Trace plots
ggplot(mcmc_long, aes(x = iteration, y = Value, color = Parameter)) +
  geom_line() +
  scale_color_manual(values = c("blue", "green", "red", "yellow")) +
  facet_wrap(~ Parameter, scales = "free_y") +
  labs(title = "Trace Plots of MCMC Samples", x = "Iteration", y = "Parameter Value") +
  theme_bw(base_size = 12)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        legend.position = "none")

# Autocorrelation and crosscorrelation plots
crosscorr.plot(out)
autocorr.plot(out)

```


